---
title: "motus_receiver_data"
author: "Patrick D. lorch"
date: "2023-08-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before running this code you should run the motus_project_data to get the data downloaded into local .motus database files.  Then you can just connect to them here.

### Signal Strength Plots

Have a look at detection data colored by antenna direction.  NA usually is for an omni.

```{r setup} 
library(motus)
library(lubridate)
library(dplyr)
Sys.setenv(TZ = "UTC")

# Project
# California Department of Fish and Wildlife Motus (#446)
proj.num = 446

## Connect to existing .motus
proj.sql.motus <- tagme(projRecv = proj.num, new = FALSE, update = FALSE)

# Focal receiver (from df.projRecvs in rectag block below)
#    stationID        stationName            serno
# 1      10593     Grizzly Island CTT-4C1B6DEFDC5C
# 2      10616        Yolo Bypass CTT-61546A685C2C
# 3      10639       Eden Landing CTT-E56F9EF340B6
# 4      12431     Elkhorn Slough CTT-3D35D88BF546
# 5      12459        Salt Slough CTT-B9683CBEE57C
# 6      12617         Llano Seco CTT-D818FFEDB922
# 7      12723        San Jacinto CTT-1CBEBDA79D53
# 8      13040         Palo Verde CTT-39A7525A06F7
# 9      13041           Imperial CTT-8B07B0305758
# 10     13310          Camp Cady CTT-V30B0154E4B7
# 11     13387 Indian Joe Springs             <NA>

focal_rec_id = 13310 # Use a stationID

## Recievers
# Set which stations data to use
# Camp Cady
rec_id_camp_cady = "CTT-V30B0154E4B7"
# Indian Joe
# rec_id_indian_joe = "CTT-V30B0154C892"

```

## Download one receiver to look at one tag

tbl.recvDeps includes all receiver deployments in the motus system, active and inactive, as of the time you updated the project .motus file. I get the table of receivers for the project in a particular way so that if Sensorstations get updated, this code still works.

```{r rectag}
# Construct receiver IDs
# To find station indentifiers, use this:
tbl.recvDeps <- tbl(proj.sql.motus, "recvDeps")
df.projRecvs <- tbl.recvDeps %>%
  filter(projectID == proj.num, # Get just CDFW receivers
         status == "active") %>%
  collect() %>%
  as.data.frame() %>%
  mutate(tsStart = as_datetime(tsStart, tz = "UTC", origin = "1970-01-01"),
         tsEnd = as_datetime(tsEnd, tz = "UTC", origin = "1970-01-01"))
View(select(df.projRecvs, c("stationID", "stationName", "serno")))

rec_id = df.projRecvs$serno[df.projRecvs$stationID == focal_rec_id]

# *** First time only ***
# These will be saved in the working directory at SQLite databases with names
#  like 'CTT-XXXXXXXXXXX.motus'
# This can take a long time
# rec.motus = tagme(rec_id, new=TRUE, update = T)

# To update the .motus SQLite file, use this
# rec.motus = tagme(rec_id, new=F, update = T)

# To just connect and use the current .motus SQLite files use this
rec.motus = tagme(rec_id, new=F, update = F)

# Get link to table
tbl.alltags <- tbl(rec.motus, "alltags")
# Bring data in from .motus file as data.frame and fix datetime
df.alltags <- tbl.alltags %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

motusLogout()
```

## Signal strength plots

```{r signal}
# Get link to table
tbl.rec.alltags <- tbl(rec.motus, "alltags")

# Bring data in from .motus file as data.frame and fix datetime
df.rec.alltags <- tbl.rec.alltags %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

(range(df.rec.alltags$ts))
which(is.na(df.rec.alltags$ts))

# ***Got to here***

# tag_motus_id = 56747 # Cowbird tag

# Filter down to tag detection time range
# *** This will fail if gps and station are not both using UTC ***
# Should/can we add a way to correct times, if needed?
# Could add multiple tags to the filter here and color by them later
# Filter by a time range created like above
# (time_range = range(points@data$timestamp))
# Whole time range. Determined from running filter step below and finding range
(time_range = c("2022-08-24", "2023-08-19"))

# YBCU tags detected by CDFW stations
# Hummus
tag_motus_id = 60762 
tag = "Hummus (60762)"
file_name_prefix = "Yellow-billed Cuckoo Hummus"
(time_range = c("2023-08-09", "2023-08-09"))

# WETA
tag_motus_id = 76180 
tag = "Western Tanager (2nd tag 76180)"
file_name_prefix = "Western Tanager 1"
(time_range = c("2023-08-12", "2023-08-12"))


# Run from here for plots
df.rec_tags = df.alltags %>%
  filter(motusTagID == tag_motus_id,
         ts >= time_range[1],
         ts <= time_range[2])
(time_range = range(df.rec_tags$ts))
# Create scaling factor for later (between 0-1)
df.rec_tags$sig_scaled = (df.rec_tags$sig
  - max(df.rec_tags$sig)) / 
  (max(df.rec_tags$sig) - min(df.rec_tags$sig))

rec_tags.sig  = df.rec_tags %>% 
  ggplot(aes(x = ts, y = sig, col = as.factor(paste(recvDeployName,
                                                    antBearing, 
                                                    sep = "-")))) +
  theme_bw() + 
  geom_point() + 
  labs(x = "Time of day (UTC)", y = "Signal strength",
       title = tag) +
  scale_color_discrete(name = "Antenna bearing") +
  facet_grid(recvDeployName ~ .)
rec_tags.sig

ggsave(filename = paste(
  file_name_prefix, tag,
  as.character(as.Date(time_range[1])),
  "to",
  as.character(as.Date(time_range[2])),
  ".pdf", sep = "_"),
  device = "pdf",
  width = 8, height = 5,
  units = "in")

ggsave(filename = paste(
  file_name_prefix, tag,
  as.character(as.Date(time_range[1])),
  "to",
  as.character(as.Date(time_range[2])),
  ".jpg", sep = "_"),
  device = "jpg",
  width = 8, height = 5,
  units = "in")

```

