---
title: "SSRS_motus_data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MotusRBook

Based on examples in **Motus R Book**
by Tara L. Crewe, Zoe Crysler, and Philip Taylor
https://beta.motus.org/MotusRBook/


```{r setup}
# install.packages("motus", 
#                  repos = c(birdscanada = 'https://birdscanada.r-universe.dev',
#                            CRAN = 'https://cloud.r-project.org'))
library(motus)
library(lubridate)
library(dplyr)
Sys.setenv(TZ = "UTC")

# Kern River Valley Tri-colored Blackbird Study
# Renamed to Southern Sierra Research Station Kern River Project
# Stopped using old name to make things more clear and eliminate mismatches
# proj.num = 247 

# rth, canebrake, migrant corner, 
# frog spring, tunnel spring, freeman canyon spring
# upper caliente creek

df.247.serno = c("CTT-957865AC5BF7", "CTT-27532C924059", "CTT-1F618DF06116", 
                 "CTT-7406F10DF482", "CTT-85E9238B9814","CTT-89E107C9320E",
                 "CTT-3A2F497F6F83")
# Uninstalled "CTT-C0612F7F986B",

# Project
## First time
proj.247.sql.motus <- tagme(projRecv = 247, new = TRUE, update = TRUE)

# Receivers

motusLogout()
```

## Update files

You can update individual sql files or all at once.

I have stopped updating the project or all at once.

```{r update}
# Project
## Update existing project from motus database
proj.247.sql.motus <- tagme(projRecv = 247, new = FALSE, update = TRUE)

# Kern River Valley Tri-colored Blackbird Study
rec.247.names = paste0(c("rth", "canebrake", "migrant corner", 
                         "frog spring", "tunnel spring", 
                         "freeman canyon spring", "upper caliente creek"), 
                       ".sql.motus")

# Individually
# Do for SSRS stations
# Update as stations are added
rec.247.names[1] <- tagme(projRecv = df.247.serno[1], new = TRUE, update = TRUE)
rec.247.names[2] <- tagme(projRecv = df.247.serno[2], new = TRUE, update = TRUE)
rec.247.names[3] <- tagme(projRecv = df.247.serno[3], new = TRUE, update = TRUE)
rec.247.names[4] <- tagme(projRecv = df.247.serno[4], new = TRUE, update = TRUE)
rec.247.names[5] <- tagme(projRecv = df.247.serno[5], new = TRUE, update = TRUE)
rec.247.names[6] <- tagme(projRecv = df.247.serno[6], new = TRUE, update = TRUE)
rec.247.names[7] <- tagme(projRecv = df.247.serno[7], new = TRUE, update = TRUE)

# All at once with variable assignment so it is comparable to the individual way
# SSRS
for (k in 1:length(df.247.serno)) {
  assign(rec.247.names[k], tagme(df.247.serno[k], update = TRUE))
}

# All at once without varialbe assignment
# for (k in 1:length(df.serno)) {
#   tagme(df.serno[k], update = TRUE)
# }

# All at once for all of your projects
### Beware this will take a really long time
# tagme()

# This will get data on all projects and receiver deployments
## Access this in recDeps table in proj.projnum.sql.motus
metadata(proj.247.sql.motus, projectIDs = 247)

motusLogout()
```

## Download one receiver to look at one tag

```{r rectag}
# install.packages("motus", 
#                  repos = c(birdscanada = 'https://birdscanada.r-universe.dev',
#                            CRAN = 'https://cloud.r-project.org'))
library(motus)
library(lubridate)
library(dplyr)
Sys.setenv(TZ = "UTC")

# Migrant corner
rec_id_migrantcorner =  "CTT-1F618DF06116"
# Red tail Hill (RTH)
rec_id_redtailhill =  "CTT-957865AC5BF7"
# Canebrake
rec_id_canebrake = "CTT-27532C924059"
# MDLT
# Tunnel Spring
rec_id_tunnel = "CTT-85E9238B9814"
# Frog Spring
rec_id_frog = "CTT-7406F10DF482"
# Freeman Canyon Spring
rec_id_freeman = "CTT-89E107C9320E"
# Upper Caliente Creek
rec_id_caliente = "CTT-3A2F497F6F83"

# Change this
# rec_id = rec_id_migrantcorner
# rec_id = rec_id_redtailhill
# rec_id = rec_id_tunnel
# rec_id = rec_id_frog
# rec_id = rec_id_canebrake
  # rec_id = rec_id_freeman
  rec_id = rec_id_caliente

# *** First time only ***
# This can take a long time
# rec.motus = tagme(rec_id, new=TRUE, update = T)

# To update the .motus SQLite file, use this
# rec.motus = tagme(rec_id, new=F, update = T)

# To just connect and use the current .motus SQLite files use this
# rec.motus = tagme(rec_id, new=F, update = F)

# Get link to table
tbl.alltags <- tbl(rec.motus, "alltags")
# Bring data in from .motus file as data.frame and fix datetime
df.alltags <- tbl.alltags %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

motusLogout()
```


## Accessing tables

Once you have the .motus files downloaded and updated, you can look at them 
using a database viewer like dBeaver by connecting an sQLite connection to the 
location on disk.

```{r data}
library(DBI)
library(RSQLite)

file.name <- dbConnect(SQLite(), "project-247.motus")
dbListTables(file.name)
dbListFields(file.name, "species")

# Kern River Valley Tri-colored Blackbird Study
# proj.num = 247
# This just points to the database in the default directory

proj.247.sql.motus = tagme(247, update = FALSE)
# virtual table
tbl.247.alltags <- tbl(proj.247.sql.motus, "alltags")
# Filter the data
# This is a filtered table with over 3 million observations
tbl.247.TRBL <- tbl(proj.247.sql.motus, "alltags") %>%
  filter(speciesEN == "Tricolored Blackbird") %>%
  collect()

# Get the daily median location of GPS points for these data
TRBL.GPS <- getGPS(src = proj.247.sql.motus, data = tbl.247.TRBL)

# Get flat table
# df.247.alltags <- tbl.247.alltags %>% 
#   collect() %>% 
#   as.data.frame()

# or
# Get table and fix datetimes
df.247.alltags <- tbl.247.alltags %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

# Remove old variables without project numbers in them below here

df.alltags.247.01_22 = df.247.alltags
# export as RDS to preserve dateTimes
# You can then just import this to work with tag data from before the date of export
saveRDS(df.alltags.247.01_22, "df_alltags_247.01_22.rds")
# To read it back in
# df.alltags.247.01_22 <- readRDS("df_alltags_247.01_22.rds")
```

## Tag Deployments


```{r tagdeploys}
# 1. download full tag metadata for our project only
# See update section above

# 2. determine how many tags are registered to your project
tbl.247.tags <- tbl(proj.247.sql.motus, "tags")
df.247.tags <- tbl.247.tags %>%
  filter(projectID == 247) %>%
  collect() %>%
  as.data.frame()
nrow(df.247.tags)
unique(df.247.tags$tagID)

# 3. determine how many of those registered tags were deployed
tbl.247.tagDeps <- tbl(proj.247.sql.motus, "tagDeps") 

df.247.tagDeps <- tbl.247.tagDeps %>%
  filter(projectID == 247) %>%
  collect() %>%
  as.data.frame() %>% # once in df format, can format dates with lubridate
  mutate(tsStart = as_datetime(tsStart, tz = "UTC", origin = "1970-01-01"),
         tsEnd = as_datetime(tsEnd, tz = "UTC", origin = "1970-01-01")) 
  # find tags that are not deployed
anti_join(df.247.tags, df.247.tagDeps, by = "tagID")
  # - currently there are 20, and some show as deployed, so not sure what this means

  # find multiple deployments
df.alltags.247.01_22 %>%
  select(motusTagID, tagDeployID) %>%
  filter(!(is.na(tagDeployID))) %>% # remove NA tagDeployIDs
  distinct() %>%
  group_by(motusTagID) %>%
  mutate(n = n()) %>%
  filter(n > 1)
  # - Currently there are 22 tags with multiple deployments
df.alltags.247.01_22 <- df.alltags.247.01_22 %>%
  mutate(motusTagDepID = paste(motusTagID, tagDeployID, sep = "."))
df.247.tagDeps <- df.247.tagDeps %>%
  mutate(motusTagDepID = paste(tagID, deployID, sep = "."))


# Plot timelines
# install.packages("vistime")
library(tidyverse)
library(ggplot2)
library(vistime)
df.247.timeline = df.247.tagDeps %>%
  select(fullID, start = tsStart, end = tsEnd) %>%
  mutate(event = str_extract(fullID, "(?<=#).+(?=:)")) %>%
  filter(!is.na(start))

p = gg_vistime(df.247.timeline)
p

# 4. determine location of tag deployments
library(ggplot2)
library(rworldmap)
na.lakes <- map_data(map = "lakes")
na.lakes <- mutate(na.lakes, long = long - 360)

# Include all of the Americas to begin
na.map <- map_data(map = "world2")
na.map <- filter(na.map, region %in% c("Canada", "USA"))
                               
na.map <- mutate(na.map, long = long- 360)

# set limits to map based on locations of detections, ensuring they include the
# deployment locations
#  These can be set arbitrarily using coordinates from Google Earth
xmin <- min(df.247.tagDeps$longitude, na.rm = TRUE) - 5 # -120 #
xmax <- max(df.247.tagDeps$longitude, na.rm = TRUE) + 5 # -116 #
ymin <- min(df.247.tagDeps$latitude, na.rm = TRUE) - 5 # 35 #
ymax <- max(df.247.tagDeps$latitude, na.rm = TRUE) + 5 # 37 #
                
# map using ggplot
ggplot(data = na.lakes, aes(x = long, y = lat)) + 
  geom_polygon(data = na.map, aes(long, lat, group = group), 
               colour = "grey", fill="grey98") + 
  geom_polygon(aes(group = group), colour = "grey", fill = "white") +
  coord_map(projection = "mercator", 
            xlim = c(xmin, xmax), 
            ylim = c(ymin, ymax)) +
  labs(x = "", y = "") + 
  theme_bw() + 
  geom_point(data = filter(df.247.tagDeps, projectID == 247), 
             aes(longitude, latitude), size = 2, shape = 1, colour = "red")
  # - One tag is deployed way out of range

# 5. determine completeness and accuracy of tag deployment metadata
  # Range of metadata values
df.247.tagDeps %>%
  select(tagID, projectID, tsStart, tsEnd, speciesID, latitude, longitude) %>%
  summary()
#  There are 
#    - two missing tsStart and end dates
#    - 12 missing species IDs
#    - 25 missing lat/lons
#    - deployment start dates are 2019-07-17 to 2021-10-13

  # Check that species IDs are appropriate for your data
sp.list <- unique(df.247.tagDeps$speciesID)
tbl.247.species <- tbl(proj.247.sql.motus, "species") 
tbl.247.species %>%
  filter(id %in% sp.list) %>%
  collect() %>%
  as.data.frame()
  # just TRBL and YBCU

```

## Reciever Metadata


```{r recievermeta}
library(tidyverse)
# proj.247.sql.motus <- tagme(247, update = TRUE)

# 1. download full receiver metadata across the network
# This may have just been run in one of the blocks above (around line 116)
metadata(proj.247.sql.motus)

# 2. determine number of project receiver deployments
tbl.247.recvDeps <- tbl(proj.247.sql.motus, "recvDeps")
df.247.projRecvs <- tbl.247.recvDeps %>%
  filter(projectID == 247, status == "active") %>%
  collect() %>%
  as.data.frame() %>%
  mutate(tsStart = as_datetime(tsStart, tz = "UTC", origin = "1970-01-01"),
         tsEnd = as_datetime(tsEnd, tz = "UTC", origin = "1970-01-01"))

summary(df.247.projRecvs)
# - Lat and lon are reasonable
# - tsStart ranges from 2019-03-22 to 2021-07-27

  # - Red Tail Hill is no longer missing
  # - missing elevation
  # - missing utcoffset

df.247.projRecvs %>%
  mutate(dateStart = date(tsStart)) %>% 
  select(-serno,-fixtureType, -macAddress, -tsStart, -tsEnd, -elevation, 
         -projectID, -status, -receiverType, -siteName) %>%
  arrange(deviceID, latitude, dateStart)

# 3. determine timing of project receiver deployments
df.247.projRecvs.long <- df.247.projRecvs %>%
  select(deviceID, deployID, tsStart, tsEnd) %>% 
  gather(when, ts, c(tsStart, tsEnd)) %>%
  # fake end date:
  mutate(ts = if_else(is.na(ts), max(ts, na.rm = TRUE) + duration(1, "month"), ts)) 

ggplot(data = df.247.projRecvs.long, 
       aes(x = ts, y = as.factor(deviceID), colour = as.factor(deployID))) +
  theme(legend.position = "none") +
  geom_line(lwd = 3) + 
  # instead, centre to the right
  geom_text(data = filter(df.247.projRecvs.long, when == "tsStart"), 
            aes(label = deployID), hjust = "left", nudge_y = 0.2, size = 3, angle = 45) +
  theme_bw() +
  labs(x = "Year", y = "Receiver ID")

ggplot(data = df.247.projRecvs.long, 
       aes(x = yday(ts), y = as.factor(deviceID), colour = as.factor(deployID))) +
  theme_bw() +
  theme(legend.position = "none") + 
  geom_line(lwd = 3) + 
  # centre labels to the left
  geom_text(data = filter(df.247.projRecvs.long, when == "tsStart"), 
            aes(label = deployID), hjust = "left", nudge_y = 0.4, size = 3) +
  labs(x = "Day of year", y = "Receiver ID") +
  facet_grid(year(ts) ~ ., scales = "free")

# 4. determine location of network-wide and project receiver deployments
df.247.recvDeps <- tbl.247.recvDeps %>%
  collect() %>%
  as.data.frame() %>%
  mutate(tsStart = as_datetime(tsStart, tz = "UTC", origin = "1970-01-01"),
         tsEnd = as_datetime(tsEnd, tz = "UTC", origin = "1970-01-01"))

na.lakes <- map_data(map = "lakes")
na.lakes <- mutate(na.lakes, long = long - 360)

na.map <- map_data(map = "world2")
na.map <- filter(na.map, 
                 region %in% c("Canada", "USA", "Mexico", "lakes", "Belize", 
                               "Costa Rica", "Panama", "Guatemala", "Honduras", 
                               "Nicaragua", "El Salvador", "Colombia", "Venezuela", 
                               "Ecuador", "Peru", "Brazil", "Guyana","Suriname", 
                               "Bolivia", "French Guiana", "Jamaica", "Cuba", 
                               "Haiti", "Dominican Republic", "The Bahamas", 
                               "Turks and Caicos Islands", "Puerto Rico", 
                               "British Virgin Islands", "Montserrat", "Dominica", 
                               "Saint Lucia", "Barbados", "Grenada", 
                               "Trinidad and Tobago", "Chile", "Argentina", 
                               "Uruguay", "Paraguay")) %>%
  mutate(long = long - 360)

xmin <- min(df.247.recvDeps$longitude, na.rm = TRUE) - 2
xmax <- -20 # restrict to the Americas (excluding a few points in Europe)
ymin <- -60 #min(df.247.recvDeps$longitude, na.rm = TRUE) - 2
ymax <- max(df.247.recvDeps$latitude, na.rm = TRUE) + 2

ggplot(data = na.lakes, aes(x = long, y = lat)) + 
  theme_bw() + 
  geom_polygon(data = na.map, aes(long, lat, group = group), 
               colour = "grey", fill = "grey98") +
  geom_polygon(aes(group = group), colour = "grey", fill = "white") +
  coord_map(projection = "mercator", xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  labs(x = "", y = "") + 
  geom_point(data = df.247.recvDeps, 
             aes(longitude, latitude, colour = as.logical(projectID == 247)), 
             size = 1, shape = 4) +
  scale_colour_manual(values = c("grey30", "red"), name = "Project 247 Deployment") 

xmin <- min(df.247.projRecvs$longitude, na.rm = TRUE) - 2
xmax <- max(df.247.projRecvs$longitude, na.rm = TRUE) + 2
ymin <- min(df.247.projRecvs$latitude, na.rm = TRUE) - 1
ymax <- max(df.247.projRecvs$latitude, na.rm = TRUE) + 1
                
# map
ggplot(data = na.lakes, aes(x = long, y = lat))+ 
  theme_bw() + 
  geom_polygon(data = na.map, aes(long, lat, group = group), 
               colour = "grey", fill = "grey98") +
  geom_polygon(aes(group = group), colour = "grey", fill = "white") +
  coord_map(projection = "mercator", xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  labs(x = "", y = "") +
  geom_point(data = filter(df.247.projRecvs, 
                           year(tsStart) %in% 2020:2021, 
                           !is.na(latitude)),  # remove mobile receivers
             aes(longitude, latitude, colour = as.factor(deviceID)), size = 2, shape = 1)+
  scale_colour_discrete(name = "Receiver ID") 


# 5. determine completeness and accuracy of receiver metadata

tbl.247.antDeps <- tbl(proj.247.sql.motus, "antDeps") 

df.247.antDeps <- tbl.247.antDeps %>%
  select(deployID, port, antennaType, bearing, heightMeters) %>%
  collect() %>%
  as.data.frame()

# receiver deployments; select variables of interest
df.247.ant.recvDeps <- df.247.recvDeps %>%
  select(deployID, receiverType, deviceID, name, latitude, longitude, 
         isMobile, tsStart, tsEnd, projectID, elevation) 

df.stationDeps <- left_join(df.247.ant.recvDeps, df.247.antDeps, by = "deployID")

# Filter to just our project if needed
df.247.stationDeps <- filter(df.stationDeps, projectID == 247)

summary(df.stationDeps)
summary(df.247.stationDeps)
# Our station deployments seem reasonable
# Current deployments
#  Red Tail Hill 2020-10-02
#  Long Canyon   2020-10-21
#  Canebrake     2021-07-27

```


## Data cleaning

```{r cleaning}
Sys.setenv(tz = "UTC")

library(motus)
library(tidyverse)
library(lubridate)

# Kern River Valley Tri-colored Blackbird Study
# proj.num = 247 

# proj.247.sql.motus <- tagme(247, update = F)

# tbl(proj.247.sql.motus, "alltags") %>% # This brings data in from local database
df.alltags.247.03_21 %>% # This uses the local copy created above
  filter(tagProjID == 247) %>% # subset to include only tags registered to project
  count(motusTagDepID) %>%
  as.data.frame()

# tbl(proj.247.sql.motus, "alltags") %>% # This brings data in from local database
df.alltags.247.01_22 %>% # This uses the local copy created above
  filter(tagProjID == 247) %>% # subset to include only tags registered to project
  mutate(rl.gt.3 = if_else(runLen == 3, "run 3", "run > 3")) %>%
  count(motusTagDepID, rl.gt.3) %>%
  collect() %>%
  spread(key = rl.gt.3, value = n)
# - large numbers have NA for deployment ID
# - some of these have very large numbers of short runs

filter(tbl(proj.247.sql.motus, "alltags"), runLen <= 3) %>% 
  collect() %>% 
  nrow()
# - over 303,000 short runs

to_remove <- tbl(proj.247.sql.motus, "runs") %>%
  select(runID, motusFilter) %>%
  filter(motusFilter == 0)
tbl_filtered <- anti_join(tbl(proj.247.sql.motus, "alltags"), to_remove, by = "runID")

filter(tbl_filtered, runLen <= 3) %>% 
  collect() %>% 
  nrow()
# 9197 rows have runLen <=3

tbl.247.filtered.alltags <- filterByActivity(proj.247.sql.motus, return = "all")

# Filter out the 9197 rows of short run records

# getGPS expects ts to not be transformed, I think, so commenting out last part
df.247.alltags.sub <- tbl.247.filtered.alltags %>% 
  filter(probability == 1) %>%
  collect() %>%
  as.data.frame() #%>%
  # mutate(ts = as_datetime(ts),  # work with dates AFTER transforming to flat file
  #        tagDeployStart = as_datetime(tagDeployStart),
  #        tagDeployEnd = as_datetime(tagDeployEnd))

df.247.block.0 <- filter(tbl.247.filtered.alltags, probability == 0) %>%
  select(motusTagID, runID) %>%
  distinct() %>%
  collect() %>%
  data.frame()

# Get GPS data
# Retrieve GPS data for each hitID
# This failed with an error
#  When you use `by = 'closest'`
# Error: Problem with `mutate()` column `gpsID`.
# i `gpsID = purrr::map_int(...)`.
# x can only subtract from "POSIXt" objects
# Run `rlang::last_error()` to see where the error occurred.
#  when you use `by = 'daily'`, which I think I used before
# Error: Problem with `mutate()` column `timeBin`.
# i `timeBin = as.integer(.data$ts/by)`.
# x '/' not defined for "POSIXt" objects
# Run `rlang::last_error()` to see where the error occurred.
# This error is due to not running the step above that transforms ts into dateTime

# Use only one of the options below, in order of likely speed
gps_index.247 <- getGPS(proj.247.sql.motus, data = df.247.alltags.sub, by = "daily")
gps_index.247 <- getGPS(proj.247.sql.motus, data = df.247.alltags.sub, by = "closest", cutoff = 20)
gps_index.247 <- getGPS(proj.247.sql.motus, data = df.247.alltags.sub, by = 15)


# Merge GPS points in with our data
df.alltags.247.sub <- left_join(df.247.alltags.sub, gps_index.247, by = "hitID")

filter(df.alltags.247.sub, is.na(gpsLat)) %>% 
  collect() %>% 
  nrow()
filter(df.alltags.247.sub, is.na(gpsLat) & is.na(recvDeployLat)) %>% 
  collect() %>% 
  nrow()
filter(df.alltags.247.sub, is.na(0)) %>% 
  collect() %>% 
  nrow()
filter(df.alltags.247.sub, is.na(999)) %>% 
  collect() %>% 
  nrow()

df.alltags.247.sub.2 <- df.alltags.247.sub %>% 
  mutate(recvLat = if_else((is.na(gpsLat)|gpsLat == 0|gpsLat == 999), 
                           recvDeployLat, gpsLat),
         recvLon = if_else((is.na(gpsLon)|gpsLon == 0|gpsLon == 999), 
                           recvDeployLon, gpsLon),
         recvAlt = if_else(is.na(gpsAlt), recvDeployAlt, gpsAlt)) %>%
  select(-noise, -slop, -burstSlop, -done, -bootnum, -mfgID, 
         -codeSet, -mfg, -nomFreq, -markerNumber, -markerType, 
         -tagDepComments, -fullID, -deviceID, -recvDeployLat, 
         -recvDeployLon, -recvDeployAlt, -speciesGroup, -gpsLat, 
         -gpsLon, -recvAlt, -recvSiteName) %>%
  mutate(recvLat = plyr::round_any(recvLat, 0.05), 
         recvLon = plyr::round_any(recvLon, 0.05),
         recvDeployName = if_else(is.na(recvDeployName), 
                                  paste(recvLat, recvLon, sep=":"), 
                                  recvDeployName))

df.alltags.247.sub.2 %>%
  filter(is.na(recvLat)) %>%
  select(recvLat, recvLon, recvDeployName, recvDeployID, recv, 
         recvProjID, recvProjName) %>%
  distinct()
# None missing on 2022-01-19

fun.getpath <- function(df, pr.nm = 247) {
  df %>%
    filter(tagProjID == pr.nm, # keep only tags registered to the sample project
           !is.na(recvLat) | !(recvLat == 0)) %>% # drops data without lon/lat
    group_by(motusTagID, runID, recvDeployName, ambigID, 
             tagDepLon, tagDepLat, recvLat, recvLon) %>%
    #summarizing by runID to get max run length and mean time stamp:
    summarize(max.runLen = max(runLen), ts.h = mean(lubridate::as_datetime(ts))) %>% 
    arrange(motusTagID, ts.h)
} # end of function

df.alltags.247.path <- fun.getpath(df.alltags.247.sub.2)

df.alltags.247.path %>%
  select(tagDepLon, tagDepLat, recvLat, recvLon, max.runLen, ts.h) %>%
  summary()


tags_all = unique(df.alltags.247.path$motusTagID)
length(unique(df.alltags.247.path$motusTagID))/6

tagset1 = tags_all[1:6]
tagset2 = tags_all[7:12]
tagset3 = tags_all[13:18]
tagset4 = tags_all[19:24]
tagset5 = tags_all[25:30]
tagset6 = tags_all[31:36]
tagset7 = tags_all[37:42]
tagset8 = tags_all[43:48]
tagset9 = tags_all[49:54]
tagset10 = tags_all[55:60]
tagset11 = tags_all[61:66]
tagset12 = tags_all[67:72]
tagset13 = tags_all[73:75]

fun.plot_path = function(path, tagset) {
  ggplot(data = filter(path, 
                     motusTagID %in% tagset), 
       aes(x = ts.h, y = recvLat)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  geom_point() + 
  geom_path() +
  facet_wrap(~ motusTagID, scales = "free", ncol = 2)
}

fun.plot_path(df.alltags.247.path, tagset1)
fun.plot_path(df.alltags.247.path, tagset2)
fun.plot_path(df.alltags.247.path, tagset3)
fun.plot_path(df.alltags.247.path, tagset4)
fun.plot_path(df.alltags.247.path, tagset5)
fun.plot_path(df.alltags.247.path, tagset6)
fun.plot_path(df.alltags.247.path, tagset7)
fun.plot_path(df.alltags.247.path, tagset8)
fun.plot_path(df.alltags.247.path, tagset9)
fun.plot_path(df.alltags.247.path, tagset10)
fun.plot_path(df.alltags.247.path, tagset11)
fun.plot_path(df.alltags.247.path, tagset12)
fun.plot_path(df.alltags.247.path, tagset13)



```


### Signal Strength Plots

Have a look at detection data colored by antenna direction.  NA usually is for an omni.

```{r sig} 
library(lubridate)
library(ggplot2)

# Set which stations data to use
# Migrant corner
rec_id_migrantcorner =  "CTT-1F618DF06116"
# Red tail Hill (RTH)
rec_id_redtailhill =  "CTT-957865AC5BF7"
# Canebrake
rec_id_canebrake = "CTT-27532C924059"
# MDLT
# Tunnel Spring
rec_id_tunnel = "CTT-85E9238B9814"
# Frog Spring
rec_id_frog = "CTT-7406F10DF482"
# Freeman Canyon Spring
rec_id_freeman = "CTT-89E107C9320E"

# Just connect and use the current .motus SQLite files created or updated above (around line 108)
rec.motus_migcor = tagme(rec_id_migrantcorner, new=F, update = F)
rec.motus_rth = tagme(rec_id_redtailhill, new=F, update = F)
rec.motus_cane = tagme(rec_id_canebrake, new=F, update = F)
rec.motus_tunl = tagme(rec_id_tunnel, new=F, update = F)
rec.motus_frog = tagme(rec_id_frog, new=F, update = F)
rec.motus_freeman = tagme(rec_id_freeman, new=F, update = F)
rec.motus_caliente = tagme(rec_id_caliente, new=F, update = F)

# Get link to table
tbl.alltags_migcor <- tbl(rec.motus_migcor, "alltags")
tbl.alltags_rth <- tbl(rec.motus_rth, "alltags")
tbl.alltags_cane <- tbl(rec.motus_cane, "alltags")
tbl.alltags_tunl <- tbl(rec.motus_tunl, "alltags")
tbl.alltags_frog <- tbl(rec.motus_frog, "alltags")
tbl.alltags_frmn <- tbl(rec.motus_freeman, "alltags")
tbl.alltags_clnt <- tbl(rec.motus_caliente, "alltags")

# Bring data in from .motus file as data.frame and fix datetime
df.alltags_migcor <- tbl.alltags_migcor %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

df.alltags_rth <- tbl.alltags_rth %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

df.alltags_cane <- tbl.alltags_cane %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

df.alltags_tunl <- tbl.alltags_tunl %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

df.alltags_frog <- tbl.alltags_frog %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

df.alltags_frmn <- tbl.alltags_frmn %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

df.alltags_clnt <- tbl.alltags_clnt %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

# Change these for each run to combine dataframes for plotting on one timeline
# YBCUs
df.alltags = rbind(df.alltags_migcor, 
                   df.alltags_rth)

# Western Sandpiper 2022
df.alltags = rbind(df.alltags_migcor, 
                   df.alltags_rth,
                   df.alltags_tunl, 
                   df.alltags_frog)

# Red Knot 2023
df.alltags = rbind(df.alltags_clnt)

# Hoary Bat 2023
df.alltags = df.alltags_frmn

# Tricolored Blackbird 2022
df.alltags = rbind(df.alltags_rth,
                   df.alltags_cane)


(range(df.alltags$ts))
which(is.na(df.alltags$ts))


# tag_motus_id = 56747 # Cowbird tag

# Filter down to tag detection time range
# *** This will fail if gps and station are not both using UTC ***
# Should/can we add a way to correct times, if needed?
# Could add multiple tags to the filter here and color by them later
# Filter by a time range created like above
# (time_range = range(points@data$timestamp))
# Whole time range. Determined from running filter step below and finding range
(time_range = c("2023-07-02", "2023-07-28"))

# YBCU tags detected by SSRS stations
# Cookie
tag_motus_id = 65544 
tag = "Cookie (65544)"
(time_range = c("2022-07-02", "2022-07-17"))
#  Short time on >1 antenna
(time_range = c("2022-07-13", "2022-07-14"))

tag_motus_id = 69554 
tag = "Cookie (2nd tag 69554)"
(time_range = c("2022-08-05", "2022-08-17"))

tag_motus_id = 65557
tag = "Stroopwafel (65557)"
(time_range = c("2022-07-05", "2022-07-23"))
(time_range = c("2022-07-17", "2022-07-23"))

tag_motus_id = 56746
tag = "YBCU Stroopwafel (2nd tag 56746)"
(time_range = c("2022-08-03", "2022-08-09"))
(time_range = c("2023-07-02", "2023-07-28"))

tag_motus_id = 56726
tag = "Snickerdoodle (56726)"
(time_range = c("2022-07-23", "2022-08-15"))
(time_range = c("2022-08-12", "2022-08-14"))

tag_motus_id = 56726
tag = "Snickerdoodle (56726)"
(time_range = c("2022-07-23", "2022-08-15"))

tag_motus_id = 60762
tag = "Hummus (60762)"
(time_range = c("2023-07-20", "2023-07-28"))

# WESA tag detected locally from Pacific Shorebirds#86:13 M.64337
tag_motus_id = 64337
tag = "Western Sandpiper (64337)"
# change these
(time_range = c("2022-07-25", "2022-07-27"))

# REKN tag detected locally from NW Mexico project M.64960 and  M.60092
tag_motus_id = 64960
tag = "Red Knot (64960)"
(time_range = c("2023-05-11", "2023-05-13"))

tag_motus_id = 60092
tag = "Red Knot (60092)"
(time_range = c("2023-05-26", "2023-05-28"))

# Hoary bat tag detected locally from CDFW motus tagged in Marin county
tag_motus_id = 70412
tag = "Hoary Bat (70412)"
(time_range = c("2023-04-21", "2023-04-24"))

# TRBL tags detected locally from SSRS Kern River#261:13.1 M.54819
tag_motus_id = 54819
tag = "Tricolored Blackbird (54819)"
# change these
(time_range = c("2022-04-08", "2022-11-10"))
(time_range = c("2022-04-08", "2022-05-21"))
# Foraging
(time_range = c("2022-05-17", "2022-05-18"))
# Roosting
(time_range = c("2022-09-11", "2022-09-13"))
(time_range = c("2022-09-11", "2022-09-28"))

# Run from here for plots
df.rec_tags = df.alltags %>%
  filter(motusTagID == tag_motus_id,
         ts >= time_range[1],
         ts <= time_range[2])
(time_range = range(df.rec_tags$ts))
# Create scaling factor for later (between 0-1)
df.rec_tags$sig_scaled = (df.rec_tags$sig
  - max(df.rec_tags$sig)) / 
  (max(df.rec_tags$sig) - min(df.rec_tags$sig))

rec_tags.sig  = df.rec_tags %>% 
  ggplot(aes(x = ts, y = sig, col = as.factor(paste(recvDeployName,
                                                    antBearing, 
                                                    sep = "-")))) +
  theme_bw() + 
  geom_point() + 
  labs(x = "Time of day (UTC)", y = "Signal strength",
       title = tag) +
  scale_color_discrete(name = "Antenna bearing") +
  facet_grid(recvDeployName ~ .)
rec_tags.sig

ggsave(filename = paste(
  "Hoary_bat", tag,
  as.character(as.Date(time_range[1])),
  "to",
  as.character(as.Date(time_range[2])),
  ".pdf", sep = "_"),
  device = "pdf",
  width = 8, height = 5,
  units = "in")

ggsave(filename = paste(
  "Hoary_bat", tag,
  as.character(as.Date(time_range[1])),
  "to",
  as.character(as.Date(time_range[2])),
  ".jpg", sep = "_"),
  device = "jpg",
  width = 8, height = 5,
  units = "in")

```


## Node data

### Cuckoos

Run code above to get df.rec_tags for a bird, then come back here and run this.

```{r nodes}
library(readr)
test_ybcu_nodes <- read_csv("test_ybcu_nodes.csv")

# Node detections

table(test_ybcu_nodes$nodeNum)

ybcu_nodes = test_ybcu_nodes %>%
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"), 
         tsCorrected = as_datetime(tsCorrected, tz = "UTC", origin = "1970-01-01"),
         BirdTag = case_when(motusTagID == 65544 ~ "Cookie (65544)",
                             motusTagID == 65557 ~ "Stroopwafel (65557)",
                             motusTagID == 56726 ~ "Snickerdoodle (56726)"))
```


## Leaflet map to examine data


```{r mapnodes}
# install.packages("leaflet.extras2", "geojsonsf")
library(tidyverse)
library(leaflet)
library(leaflet.extras2)
library(sf)
library(geojsonsf)
library(plotrix)
library(htmlwidgets)

node_max = ybcu_nodes %>%
  mutate(day = floor_date(ts, 'day'),
         hour = hour(ts)) %>%
  group_by(motusTagID, day, hour) %>%
  slice(which.max(sig))


node_max_sf = st_as_sf(node_max, coords = c("node_Lon", "node_Lat"))

node_max_sf = st_jitter(node_max_sf, amount = 0.0002)

pal <- colorFactor(
  palette = "magma",
  domain = node_max_sf$BirdTag,
  na.color = "#737373")

m = leaflet(node_max_sf) %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles('Esri.WorldImagery',group = "Esri World Imagery") %>%
  addCircleMarkers(color = pal(node_max_sf$BirdTag),
                   popup =
                     as.character(node_max_sf$BirdTag)) %>%
  addLayersControl(
        baseGroups = c("OSM (default)", "Esri World Imagery"),
        position = "bottomright",
        options = layersControlOptions(collapsed = TRUE)) %>%
  addLegend(pal = pal, values = node_max_sf$BirdTag,
            position = "bottomright") %>%
  addScaleBar(
    position = c("bottomleft"),
    options = scaleBarOptions()
  ) %>%
    addTimeslider(data = node_max_sf, 
                  options = timesliderOptions(
               position = "bottomleft",
               timeAttribute = "ts",
               range = TRUE))


m

saveWidget(m, file = 
             # paste(df.rec_tags$recvSiteName[1], 
             paste(df.rec_tags$recvDeployName[1], 
                   as.character(tag_motus_id), 
                   as.character(as.Date(df.rec_tags$ts[1])),
                   "html", 
                   sep = "."))
```

