---
title: "TagTestingCDFW"
author: "Patrick D. lorch"
date: "2022-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import GPS and receiver tag data and plot

## Rmotus package install

```{r eval=F}
# install.packages("motus", 
#                  repos = c(birdscanada = 'https://birdscanada.r-universe.dev',
#                            CRAN = 'https://cloud.r-project.org'))

# If individual package fail
# install.packages("crayon")
# install.packages("assertthat")
# install.packages("cpp11")
# install.packages("jsonlite")

# If this still fails you can do this and rerun
options(timeout = max(1000, getOption("timeout")))

```

## Set receiver and file paths for a given map

The way this is set up, you set up receiver, tag, .gpx combos, then comment out 
all but one and run.

You can knit this in Rstudio to produce a report for each combo that shows 
diagnostics.

```{r set}
# Camp Cady
# rec_id = "CTT-V30B0154E4B7"
# tag_motus_id = 57234
# GPS_filepath = "CDFW sites/"

# rec_id = "CTT-V30B0154E4B7"
# tag_motus_id = 56719
# GPS_filepath = "CDFW sites/"

# Indian Joe
# rec_id = "CTT-V30B0154C892"
# tag_motus_id = 57234
# GPS_filepath = "CDFW sites/"

# rec_id = "CTT-V30B0154C892"
# tag_motus_id = 56719
# GPS_filepath = "CDFW sites/"

# How many rings showing distance from tower do you want
km_rings = 4

gpx_layer = "track_points"
p_cex = 10 # Point size multiplier

```

## MotusRBook stuff

Based on examples in **Motus R Book**
by Tara L. Crewe, Zoe Crysler, and Philip Taylor
https://beta.motus.org/MotusRBook/

If this is your first time generating a receiver .motus file, use the first one.
It will generate an SQLite file called something.motus in the file storage (memory)
area with all the receiver data in tables, including tag detections. The tbl.
command below creates a link to one table in the database in memory. Then you 
can bring the data for that table into a data.frame.

```{r tagdetections}
library(motus)
library(lubridate)
library(dplyr)
Sys.setenv(TZ = "UTC")

# *** First time only ***
# This can take a long time
# rec.motus = tagme(rec_id, new=TRUE, update = T)

# To update the .motus SQLite file, use this
# rec.motus = tagme(rec_id, new=F, update = T)

# To just connect and use the current .motus SQLite files use this
rec.motus = tagme(rec_id, new=F, update = F)

# Get link to table
tbl.alltags <- tbl(rec.motus, "alltags")
# Bring data in from .motus file as data.frame and fix datetime
df.alltags <- tbl.alltags %>% 
  collect() %>% 
  as.data.frame() %>%     # for all fields in the df (data frame)
  mutate(ts = as_datetime(ts, tz = "UTC", origin = "1970-01-01"))

motusLogout()
```

## Import, reformat, and merge GPS and reciever tag detection data

Not clear how best to match GPS points to tag detections.  For now we use nearest point in time. 
Uses code example here:
https://stackoverflow.com/questions/70974563/r-merge-two-data-frames-based-on-nearest-date-and-time-match

* This file is set up to do the import and map for a single tag.
* Assumes both .gpx and sensorstation have accurate time in UTC

```{r gpsdata}
library(rgdal)
library(lubridate)
library(ggplot2)
library(dplyr)
library(motus)
library(sp)

# Check structure of .gpx file (different between devices)
ogrListLayers(GPS_filepath)

# Read in GPS points from route in .gpx format
points = readOGR(GPS_filepath, gpx_layer)
View(points@data)
# This step can be commented out if just GPS track of interest was exported
# Subset just one track segment for the trip of interest
#  Don't use date to subset, due to UTC being a day ahead during some of track
# points = points[points@data$track_fid == 63,] # For Elkhorn slough

# Create datetime from end and store in blank timestamp field and show range
# Assumes this file is already trimmed to roughly the test period
points@data = points@data %>%
  mutate(timestamp = as_datetime(time, tz = "UTC"))

(time_range = range(points@data$timestamp))

```

### Signal Strength Plots

Have a look at detection data colored by antenna direction.  NA usually is for an omni.

```{r sig} 
# Filter down to test tag and times of test
# *** This will fail if gps and station are not both using UTC ***
# Should/can we add a way to correct times, if needed?
# Could add multiple tags to the filter here and color by them later
df.rec_tags = df.alltags %>%
  filter(motusTagID == tag_motus_id,
         ts >= time_range[1],
         ts <= time_range[2])
# Create scaling factor for later (between 0-1)
df.rec_tags$sig_scaled = (df.rec_tags$sig
  - max(df.rec_tags$sig)) / 
  (max(df.rec_tags$sig) - min(df.rec_tags$sig))

rec_tags.sig  = df.rec_tags %>% 
  ggplot(aes(x = ts, y = sig, 
             col = as.factor(antBearing))) +
  theme_bw() + 
  geom_point() + 
  labs(x = "Time of day", y = "Signal strength") +
  scale_color_discrete(name = "Antenna bearing") +
  facet_grid(recvDeployName ~ .)
rec_tags.sig

```


### Match positions with detections

This is where it is not exactly clear the right way to do this.

Using minutes there are ~11-12 detections/minute and 12 GPS coordinates per minute. 
Could consider doing seconds instead.

```{r merge}
# Round to nearest minute
points@data$tmp = points@data$timestamp %>%
  round_date("minute")
df.rec_tags$tmp = df.rec_tags$ts %>%
  round_date("minute")
## Add group to GPS coords
points@data$grp = 1:nrow(points@data)

tags_points_l_joined = df.rec_tags %>%
  left_join(points@data, by = "tmp")
tags_points.nearest = tags_points_l_joined %>%
  group_by(grp) %>% 
  slice(which.min(abs(difftime(timestamp, ts)))) %>% 
  ungroup() %>% 
  select("track_seg_point_id", "ele", "timestamp", "grp", "hitID", "runID",
         "batchID", "sig", "sig_scaled", "done", "motusTagID", "ambigID", 
         "port", "nodeNum", "runLen", "motusFilter", "bootnum", "deviceID", 
         "recvDeployID", "recvDeployLat", "recvDeployLon", "recvDeployAlt", 
         "recv", "recvDeployName" , "recvSiteName", "isRecvMobile", "recvProjID",
         "recvUtcOffset", "antType", "antBearing", "antHeight", "speciesEN",
         "speciesFR", "speciesSci", "speciesGroup", "tagProjName" , 
         "recvProjName")
points@data = points@data %>%
  left_join(tags_points.nearest, by = "grp")
plot(points)

```


## Leaflet map to examine data


```{r maptesttag}
# install.packages("leaflet.extras2", "geojsonsf")
library(dplyr)
library(leaflet)
library(leaflet.extras2)
library(sf)
library(geojsonsf)
library(plotrix)
library(htmlwidgets)

pal <- colorNumeric(
  palette = "magma",
  domain = points$sig,
  na.color = "#737373")

m = leaflet(points) %>%
  addTiles(group = "OSM (default)") %>%
  addProviderTiles('Esri.WorldImagery',group = "Esri World Imagery") %>%
  addCircles(data = station_coords, color = "green", 
             radius = seq(1, km_rings)*1000,
             fillColor = "transparent") %>%
  addCircleMarkers(color = pal(points$sig),
                   radius = ifelse(is.na(points$sig_scaled), 2, -(p_cex * points$sig_scaled)),
                   popup = ~paste(as.character(points$sig),
                                  as.character(points$motusTagID), sep = "; "), 
                   label = ~paste(as.character(points$sig),
                                  as.character(points$motusTagID), sep = "; ")) %>%
  addLayersControl(
        baseGroups = c("OSM (default)", "Esri World Imagery"),
        position = "bottomright",
        options = layersControlOptions(collapsed = TRUE)) %>%
  addLegend(pal = pal, values = points$sig) %>%
  addScaleBar(
    position = c("bottomleft"),
    options = scaleBarOptions()
  )
m
saveWidget(m, file = 
             # paste(df.rec_tags$recvSiteName[1], 
             paste(df.rec_tags$recvDeployName[1], 
                   as.character(tag_motus_id), 
                   as.character(as.Date(df.rec_tags$ts[1])),
                   "html", 
                   sep = "."))
```

